---
title: "Problem Set 1"
author: Nicholas Kim
format: html

---
Link to my [github]("https://github.com/nick-kimm/Problem-Set-1") repository

Looked up some functions and syntax online to answer some questions. Ex. Syntax of if statement inside for loop. 

##Problem 1

Part a: Loading data
```{r}
wine <- read.csv("/Users/nicholaskim/Documents/STAT 506/hw1/wine.data",header = FALSE)

wine_names <- list('Wine Type','Alchohol','Malic Acid','Ash','Alcalinity of Ash','Magnesium','Total Phenols','Flavanoids','Nonflavanoid Phenols','Proanthocyanins','Color Intensity','Hue','OD280/OD315','Proline' )

colnames(wine) <- wine_names

wine
```

Part b: Checking if count of wine type matches that noted in wine_name text file
```{r}
as.data.frame(table(wine$`Wine Type`))
```
From doc we expect to see 59 for Type 1, 71 for Type 2, and 48 for Type 3. From table above we can see it matches!

Part c: Q1. What is the correlation b/t alcohol and color intensity?
```{r}
cor(wine$Alchohol,wine$`Color Intensity`)
```
The correlation between these two variables is 54.64%

Q2. Which class has the highest correlation? Which has the lowest?
```{r}

```

Q3. What is the alcohol content of the wine with the highest color intensity?
```{r}
for (x in 1:length(wine$`Wine Type`)){
  if (wine[x,11] == max(wine$`Color Intensity`)){
    alc = wine[x,2]
  }
}
alc
```
The alcohol content of the wine w/ the highest Color Intensity is 14.34%.

Q4. What percentage of wines had a higher content of proanthocyanins compare to ash?
```{r}
count = 0

for (x in 1:length(wine$`Wine Type`)){
  if (wine[x,10] > wine[x,4]){
    count = 1+count
  }
}

count/178*100
```
The percentage of wines that had a higher content of Proanthocyanins compared to Ash was 8.4%

Part d: Create a table identifying the average value of each variable, providing one row for the overall average, and one row per class with class averages.

Finding values for Overall Averages
```{r}
overall_avgs = list()

for (x in 1:14){
  overall_avgs[x] = mean(wine[,x])
}
```

Finding averages for each Wine Type respectively
```{r}
c1_avgs = list()

for (x in 2:14){
  c1_avgs[x] = mean(wine[1:59,x])
}

c2_avgs = list()
for (x in 2:14){
  c2_avgs[x] = mean(wine[60:131,x])
}

c3_avgs = list()

for (x in 2:14){
  c3_avgs[x] = mean(wine[131:178,x])
}
```

Matrix to form data is shape we want
````{r}
matrix(data = c(overall_avgs[2:14],c1_avgs[2:14],c2_avgs[2:14],c3_avgs[2:14]),nrow=4, ncol=13,
       dimnames=list(c("Overall Avg", "Type 1 Avg", "Type 2 Avg", "Type 3 Avg"),wine_names[2:14]))
```
Part e: Carry out a series of t-tests to examine whether the level of phenols differs across the three classes. Present the R output and interpret the results.
```{r}

```
##Problem 2


